\chapter{Implementation}
\label{chap:implementation}

This chapter describes how the architecture from Chapter~\ref{chap:architecture} was turned into running code. The focus is on the parts that are specific to the hybrid balance model: the database schema rework, the on-chip data layout, the sync endpoint, and the reconciliation logic that ties everything together. Topics like screen design, navigation, or generic CRUD endpoints are mentioned only briefly because they do not say much about the system as a whole.

\section{Implementation overview}
\label{sec:impl-overview}

The system is built on three codebases that live in the same monorepo. The backend is a NestJS server in \texttt{backend/}, talking to a Postgres database through Drizzle ORM. The vendor terminal is a React Native application built with Expo, in \texttt{softpos/}, using \texttt{react-native-nfc-manager} for chip reads and writes. The thesis itself lives in \texttt{docs/}.

The whole stack uses TypeScript end to end. The data shapes that travel between the chip, the terminal, and the server are easy to get wrong, and a single set of types shared between client and server catches most of those mistakes at build time. Drizzle's schema definitions also act as the source of truth for the database, so the table layout and the application types stay in sync without a separate code generation step.

\subsection{Project layout and libraries}
\label{subsec:impl-libraries}

The backend is a standard NestJS 11 project: controllers, services, modules, with class-validator on the DTOs and Drizzle for the database. Authentication uses Better Auth with the \texttt{bearer} and \texttt{emailOTP} plugins; there is no custom JWT layer. Stripe integration goes through the official \texttt{stripe} SDK. Postgres access is through the \texttt{postgres} driver.

The terminal is an Expo app. NFC is handled by \texttt{react-native-nfc-manager}. The offline queue is stored in plain JSON files via \texttt{expo-file-system}. Connectivity changes are observed through \texttt{@react-native-community/netinfo}. Cryptography on the chip side (HKDF for key derivation, HMAC-SHA256 for chip integrity) uses \texttt{@noble/hashes}. Bearer tokens for the backend are kept in \texttt{expo-secure-store}, since React Native does not have a cookie jar. HTTP requests go through \texttt{axios} and server state is cached with \texttt{@tanstack/react-query}.

\subsection{Deployment}
\label{subsec:impl-deployment}

The backend runs on DigitalOcean's App Platform, behind a managed load balancer. The database is a serverless Neon Postgres instance in a European region, accessed through \texttt{@neondatabase/serverless} on the backend side. The choice of region is driven by the latency profile we want for festivals in Romania and the rest of Europe.

\section{Database schema for the hybrid model}
\label{sec:impl-schema}

The hybrid model from Chapter~\ref{chap:architecture} required a real cleanup of the schema. The original design had a generic \texttt{wallets} table where each user owned a single balance, shared across events. That made sense when the system was thought of as a closed-loop wallet platform, but it does not match how festivals actually work. An attendee at one festival does not expect leftover credit to appear at a different festival run by a different organizer, and an organizer wants their event balances to be isolated for accounting and refunds. Balance is therefore event-scoped, and the natural place for it is the row that already represents the link between an attendee and a specific event.

That row is \texttt{event\_bracelets}, and it now carries the balance directly. Three new columns were added: \texttt{balance} (an integer count of cents, defaulting to zero), \texttt{debit\_counter\_seen} (the highest \texttt{debit\_counter} value the server has applied from the chip for this bracelet), and \texttt{credit\_counter} (a server-owned counter that increases by one every time the server credits the bracelet, for example after a successful top-up). The two counters are independent on purpose. Debits are allocated by the chip when a payment happens at a vendor, and credits are allocated by the server when money flows in. Keeping them separate means neither side has to coordinate with the other to pick a fresh value.

The \texttt{transactions} table was reshaped to match. The legacy \texttt{transaction\_counter} column was dropped because it tried to merge debits and credits into one sequence, which made it impossible to tell which side a row was authored by. In its place, two nullable columns were added: \texttt{debit\_counter}, set on debit rows, and \texttt{credit\_counter}, set on credit rows. Exactly one of the two is populated for any given transaction. The foreign key \texttt{wallet\_id} was retargeted to \texttt{event\_bracelet\_id}, and the redundant \texttt{event\_id} column was removed because the event is reachable through the bracelet. The same retargeting was applied to \texttt{payment\_intents.wallet\_id}.

Two tables were dropped entirely. The \texttt{wallets} table is no longer needed because the balance moved to \texttt{event\_bracelets}, and \texttt{vendor\_invitations} was removed because the invitation flow was replaced by a simpler model where adding a user as a vendor creates the user account on the spot if they do not already have one. The remaining membership tables (\texttt{event\_tickets}, \texttt{event\_members}, \texttt{vendor\_members}) were kept as is.

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|}
		\hline
		\textbf{Table} & \textbf{Column} & \textbf{Purpose} \\
		\hline
		event\_bracelets & balance & Current balance in cents, event scoped \\
		\hline
		event\_bracelets & debit\_counter\_seen & Last debit counter applied from chip \\
		\hline
		event\_bracelets & credit\_counter & Server-owned credit counter \\
		\hline
		transactions & debit\_counter & Set on debit rows, nullable \\
		\hline
		transactions & credit\_counter & Set on credit rows, nullable \\
		\hline
		transactions & event\_bracelet\_id & Replaces wallet\_id \\
		\hline
		payment\_intents & event\_bracelet\_id & Replaces wallet\_id \\
		\hline
	\end{tabular}
	\caption{New and changed columns introduced by the hybrid model rework.}
	\label{tab:schema-changes}
\end{table}

The schema also relies on Postgres for two safety nets that the application code does not have to repeat. A check constraint on \texttt{event\_bracelets} requires \texttt{balance >= 0}, so any path that would push a balance negative aborts the transaction at the database layer. Two partial unique indexes scoped on \texttt{status = 'active'} make sure that, for any given event, there is at most one active assignment per wristband UID and at most one active assignment per user. These two indexes are partial on purpose: revoked or replaced rows stay around for audit, but they do not block re-issuing a bracelet to the same person.

The end result is a schema that maps cleanly onto how operators think about a festival. A bracelet belongs to an event, the balance lives on that bracelet, and every movement of money against it is logged in \texttt{transactions} with a counter that says whether it was a debit (allocated by the chip) or a credit (allocated by the server). The full table list is: \texttt{users}, \texttt{sessions}, \texttt{accounts}, \texttt{verifications} (Better Auth), \texttt{events}, \texttt{event\_members}, \texttt{vendor\_members}, \texttt{vendors}, \texttt{event\_bracelets}, \texttt{event\_tickets}, \texttt{transactions}, \texttt{payment\_intents}, and \texttt{audit\_logs}. Every table that belongs to an event is reachable from the event row, either through an \texttt{event\_id} foreign key or through \texttt{event\_bracelets}, which gives the multi-tenant boundary without a tenant column on every table.

\section{Authentication and bearer tokens}
\label{sec:impl-auth}

Authentication is handled by Better Auth on both sides. The backend declares the auth instance in \texttt{src/common/auth/auth.ts}, with the \texttt{expo}, \texttt{bearer}, and \texttt{emailOTP} plugins. There is no custom JWT layer: the server issues an opaque session and the mobile app sends the session token as \texttt{Authorization: Bearer <token>}. The token itself is kept in \texttt{expo-secure-store}, since React Native does not have a cookie jar to lean on the way a web browser does. Login goes through the email-OTP flow: the user types their email, the backend sends a six-digit code, the app posts the code back, and Better Auth returns the token. Authorization for vendor and admin operations is done through the \texttt{event\_members} and \texttt{vendor\_members} tables rather than a global role on the user row, so the same person can be an attendee at one event and a vendor operator at another without a role change at the user level.

\section{The reconciliation service and sync endpoint}
\label{sec:impl-sync}

The terminal sends the queued debits it has for one bracelet, and the server returns the outcome the terminal needs to clean up its local queue.

\begin{verbatim}
POST /api/v1/events/:eventId/wristbands/:wristbandUid/sync
Authorization: Bearer <token>
Content-Type: application/json

Request:
{
  "pendingDebits": [
    {
      "idempotencyKey": "<uuid>",
      "amount": <int cents>,
      "vendorId": "<uuid>",
      "counterValue": <int>,
      "deviceId": "<string>",
      "clientTimestamp": "<ISO 8601>"
    }
  ]
}

Response 200:
{
  "serverState": {
    "balance": <int cents>,
    "debit_counter_seen": <int>,
    "credit_counter": <int>
  },
  "applied":  ["<idempotencyKey>", ...],
  "rejected": [
    { "idempotencyKey": "...",
      "reason": "duplicate" | "insufficient_funds" }
  ],
  "chipShouldWrite": {
    "balance": <int cents>,
    "credit_counter": <int>
  }
}
\end{verbatim}

The terminal does not send the chip's HMAC, because by the time \texttt{pendingDebits} reaches the server those debits have already been written to the chip and the HMAC has been recomputed. What matters at sync time is whether each entry in the queue has actually landed on the server's ledger.

The reconciliation algorithm runs inside a serializable Postgres transaction (\texttt{SET TRANSACTION ISOLATION LEVEL SERIALIZABLE}) that locks the bracelet row for the duration of the call. Serializable isolation matters here: two terminals that both held the same bracelet during overlapping offline windows can hit the sync endpoint at the same time, and the database has to make their effects look like a strict ordering. With a weaker level a race between two reconciliations could let the same chip counter be applied twice or could allow the balance to go negative.

\begin{verbatim}
1. Open a serializable transaction and lock the bracelet row.
2. Sort pendingDebits by counterValue ascending.
3. For each pending debit:
     a. If idempotencyKey already exists in transactions,
        record it as duplicate and continue.
     b. If counterValue <= bracelet.debit_counter_seen,
        insert a historical debit if that counter is not
        already present, without changing bracelet.balance,
        then continue.
     c. If bracelet.balance < amount,
        record it as insufficient_funds and continue.
     d. Decrement bracelet.balance by amount.
     e. Set bracelet.debit_counter_seen = counterValue.
     f. Insert a transaction row with debit_counter
        = counterValue and the supplied idempotencyKey.
4. Read the current bracelet.balance and bracelet.credit_counter.
5. Build chipShouldWrite from that current state.
6. Commit the transaction.
7. Return serverState, applied, rejected, and chipShouldWrite.
\end{verbatim}

The two rejection reasons cover the cases the service produces today. The \texttt{duplicate} reason fires when the same debit reaches the server twice, either because the terminal retried after a flaky response or because a transaction row for the same bracelet counter already exists. The \texttt{insufficient\_funds} reason fires when the queued debit is ahead of the server state and the server cannot honour it from the current balance. A late debit whose counter is already behind \texttt{debit\_counter\_seen} is not rejected automatically: if no row exists for that counter, the service records it as historical itemized vendor data and marks the metadata as already absorbed. The \texttt{RejectReason} type leaves room for two more categories (\texttt{counter\_gap} for non-contiguous counters and \texttt{invalid} for malformed entries) but the current service does not produce them; they are reserved for future hardening.

The response still carries \texttt{chipShouldWrite} for observability and parity with the online charge response, but background sync does not write to the chip. Chip writes happen only during live NFC interactions, especially online charge, where the bracelet is present and the server can compare its state with the database.

\section{Chip data layout, HMAC, and the key chain}
\label{sec:impl-chip-layout}

The chips used for the bracelets are NTAG class tags, which are cheap and widely available, but do not have the cryptographic features of DESFire. They cannot do mutual authentication on their own, so the security has to come from a signed record written into the user-memory pages. The on-chip layout is small and fixed.

\begin{verbatim}
balance              uint32 little-endian (cents)
debit_counter        uint32 little-endian
credit_counter_seen  uint32 little-endian
hmac                 bytes(16)  HMAC-SHA256 truncated to 128 bits
                                over the 12 bytes above
\end{verbatim}

The first twelve bytes are the state. The last sixteen bytes are an HMAC computed over those twelve bytes using a per-bracelet symmetric key. Anyone reading the chip can read the state, but only someone who knows the per-bracelet key can produce a record that the terminal will accept.

The per-bracelet key is derived rather than stored. The terminal holds a master key and computes the bracelet's working key on the fly with HKDF-SHA256:

\begin{verbatim}
key = HKDF(masterKey, salt = wristbandUid, info = "bracelet-v1")
\end{verbatim}

The wristband's UID is read from the chip's read-only manufacturer area, so it cannot be spoofed by simply rewriting the user memory. The \texttt{info} string \texttt{"bracelet-v1"} fences the derivation against reuse if a future version of the layout ever needs to coexist with this one.

In the current MVP the master key sits on the terminal: it is read from the \texttt{EXPO\_PUBLIC\_BRACELET\_MASTER\_KEY} environment variable at build time, with a hex fallback baked into the source so a fresh checkout still runs. Expo inlines \texttt{EXPO\_PUBLIC\_*} variables into the JavaScript bundle, so anyone with the app binary can extract the master key and derive every per-bracelet key. This is acceptable for a thesis demo but it is not a production posture; Chapter~\ref{chap:security} discusses the trust implications and the upgrade path (per-event keys provisioned through an HSM, or moving to a DESFire chip that holds its own key).

\section{Online debit path}
\label{sec:impl-online-debit}

When a vendor terminal has connectivity, the simplest and safest thing to do is to push the debit to the server first and let the chip catch up afterwards. The terminal posts the debit to a charge endpoint that takes the bracelet, the amount, the vendor identity, and an idempotency key generated on the device. The server opens a serializable transaction on the bracelet row, checks that the balance covers the amount, decrements the balance, allocates the next \texttt{debit\_counter} value (one above whatever was last applied), inserts a \texttt{transactions} row with that counter, and returns the new \texttt{balance} and \texttt{debit\_counter} to the terminal.

The terminal then writes the new state back to the chip in a single NFC session, recomputing the HMAC over the new balance, debit counter, and the unchanged \texttt{credit\_counter\_seen}. If that write succeeds, the transaction is closed. If it fails because the customer pulled their wrist away too early, the server-side debit has already happened and the chip is now stale relative to the ledger. The next time the chip is read by any terminal, the server's view will be lower than the chip's view, and on the next sync the terminal can write the corrected state down. This is the same self-healing path the offline flow relies on, just hit from a different angle.

The decision to do server-first online and chip-first offline is intentional. When the network is up, the server is the cheapest place to enforce invariants and the chip is the more failure-prone half. When the network is down, the chip is the only place where the invariant can be enforced at all. The model swaps which side leads based on what is actually available.

\section{Offline debit path and the queue}
\label{sec:impl-offline-debit}

The offline path starts on the terminal. When connectivity is missing, or when the request to the charge endpoint times out, the terminal falls back to local authorisation. It reads the chip, derives the per-bracelet key from the bundled master key with HKDF, verifies the HMAC, and checks that the balance covers the amount.

If the check passes, the terminal performs the debit on the chip inside a single NFC session. The new balance is computed (\texttt{balance - amount}), the debit counter is incremented by one, the credit counter seen is left untouched, the HMAC is recomputed over the three new fields, and the resulting twenty-eight bytes are written back to the chip. The blob spans seven NTAG pages, which the transport layer writes one page at a time (NTAG WRITE is a per-page command). The multi-page write is therefore not atomic on its own. What makes the chip-side update behave atomically as far as the rest of the system is concerned is the HMAC: a torn write leaves the body bytes inconsistent with the trailing HMAC, and the next read fails verification before any logic acts on the partial state. The terminal is then expected to retry the write or fall back to the queue.

To stop concurrent NFC operations from stepping on each other, all chip reads and writes go through a single \texttt{withSession()} helper that chains them through one promise, so that the higher level code can call \texttt{readChip} and \texttt{writeChip} from anywhere without having to think about overlapping sessions.

Only once the chip write is acknowledged does the terminal enqueue the pending debit on disk. The queue is stored as plain JSON files under the app's document directory, one file per (event, vendor) pair, written through \texttt{expo-file-system}. Each entry carries the same fields the sync request will eventually need: a generated \texttt{idempotencyKey}, the \texttt{amount}, the \texttt{vendorId}, the \texttt{counterValue} that was just written to the chip, the \texttt{deviceId}, and the local timestamp. Storing the entry only after the chip write means the queue can never contain a debit that did not happen; the worst case is a chip update that the terminal forgot, which the next reconciliation will detect through the counter.

When connectivity comes back, the terminal posts the queue to the sync endpoint described in Section~\ref{sec:impl-sync}. The server replies with which entries it accepted and which it rejected. The terminal removes the accepted entries from the JSON file. Rejected entries are kept and surfaced to the operator with the reason: \texttt{duplicate} is silently dropped, and \texttt{insufficient\_funds} is escalated to a manual review screen because it implies the queued debit and the server balance disagreed in a way the automatic logic should not paper over.

\subsection{Auto-sync on reconnect}
\label{subsec:impl-autosync}

The terminal does not wait for the operator to re-tap a bracelet before flushing the queue. A subscription to \texttt{@react-native-community/netinfo} watches the connectivity state, and on a transition from offline to online it triggers a queue drain across every (event, vendor) pair the terminal has seen during the session. The drain only uploads queued debits; it does not require the bracelet to be present and does not send a stored chip snapshot. A subsequent online charge tap is responsible for bringing the chip back to the server-returned state.

\section{Top-up path and the Stripe webhook}
\label{sec:impl-topup}

Top-ups are the credit side of the system, and they always go through the server because they involve real money moving through Stripe. The terminal (or the attendee's own phone, depending on the flow) creates a Stripe Payment Intent against the backend, which returns the client secret. The user confirms the payment in the Stripe SDK. On success, Stripe sends a \texttt{payment\_intent.succeeded} webhook to the backend.

The backend has \texttt{rawBody} enabled at the application level, so the webhook controller can read \texttt{req.rawBody} and pass it together with the \texttt{Stripe-Signature} header to the Stripe SDK for signature verification. Verifying against the raw body, before any JSON parser has touched it, is the only way the signature will check out; this is why a body parser cannot be turned on globally.

Once the signature is valid, the handler looks up the corresponding \texttt{payment\_intents} row by the intent ID and runs the credit inside a Postgres transaction. The transaction does not set an isolation level, so it runs at the default \texttt{read committed} level. The handler increments \texttt{bracelet.balance} by the topped-up amount, increments \texttt{bracelet.credit\_counter} by one, and inserts a \texttt{transactions} row with that new \texttt{credit\_counter} value and a reference to the payment intent.

Idempotency for top-ups does not come from a unique key on the credit transaction. Instead, two layers protect against duplicate credits. The \texttt{payment\_intents} table has a \texttt{UNIQUE} constraint on \texttt{provider\_intent\_id}, so two webhooks for the same Stripe intent cannot create two rows. On top of that, the handler short-circuits if the intent row is already in status \texttt{succeeded}: a webhook retry from Stripe finds the intent already marked as completed, returns immediately, and does not run the credit a second time. The combination is enough to make the webhook handler safe to retry, which Stripe relies on aggressively.

After the credit is committed, the next online charge tap returns a \texttt{chipShouldWrite} that reflects the new balance and the new \texttt{credit\_counter}. The terminal writes this back to the chip, and the chip's \texttt{credit\_counter\_seen} catches up to the server's \texttt{credit\_counter}. From that point on the chip and the server agree, and the bracelet is ready to be debited again, online or offline.

\section{Tests}
\label{sec:impl-tests}

The backend ships with three test suites. \texttt{reconciliation.e2e-spec.ts} drives the sync endpoint against a real Postgres started inside a Testcontainers container, exercising idempotency, counter checks, balance underflow, and the worked example from Section~\ref{subsec:reconciliation-example}. \texttt{bracelets.e2e-spec.ts} covers bracelet assignment, replacement, and the issuance of the short-lived bracelet token used by the terminal during a session, against the same kind of containerised database. \texttt{bracelet-token.spec.ts} is a unit test that exercises the token signing, verification, and expiry logic without touching the database. The terminal does not yet have a test suite of its own; the parts of the chip handling that have a testable surface (the codec, the HKDF derivation) are covered indirectly through the backend-side end-to-end tests.
