%\documentclass[12pt]{scrreprt}
\documentclass[12pt]{report}

% language may be romanian or english (default is english)
% type may be bachelor or master (default is bachelor)
\usepackage[language=english, type=master]{style}

%\geometry{a4paper,top=2.5cm,left=3cm,right=2.5cm,bottom=2.5cm}
%in style
%controlling the appearance of your headers and footers
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\chead{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.2pt}

\begin{document}

\specialization{Sisteme informatice avansate}
\title{A Cloud-Based Closed-Loop Payment System with Hybrid On-Chip Reconciliation for Festival Environments}
\author{Tudor Esan}
\supervisor{Assoc. Prof. Christian Sacarea, PhD}

\maketitle


\newpage
\thispagestyle{empty}
\mbox{}
\newpage
\pagenumbering{roman}

\cleardoublepage
ABSTRACT
\vspace{0.5cm}
\hrule
\vspace{0.5cm}
%\cleardoublepage

This thesis describes a cloud-based closed-loop payment system for music festivals and large outdoor events. The system is online by default and uses the bracelet balance as a fallback when the network goes down. Festivals are known for bad connectivity when a lot of people gather in the same place, but in our view, building the whole system around the assumption that the network will fail brings more complexity than it is worth. We keep things online most of the time using a Starlink uplink and a redundant network topology, and we let the NFC wristband hold the balance as a cache for the rest. When the network is up, transactions go through the server in real time. When it is not, the chip itself stores the balance and a monotonic debit counter, so vendor terminals can approve a payment locally and not let anyone spend money twice. As soon as connectivity comes back, the terminals push their queued transactions to the server in batches, and the server applies them against its own credit counter. Credits are written only by the server and debits are written only by the chip, so the two flows meet again on their own when the network is back. We assess the architecture against what a real festival actually needs and we argue that treating offline as a bounded fallback, and not as the foundation of the system, gives a simpler and more maintainable result.

\tableofcontents


\newpage
\pagenumbering{arabic}

\input{chapters/chapter1_introduction}

\input{chapters/chapter2}

\input{chapters/chapter4_architecture}

%\input{chapters/chapter3_technologies} % Removed for now

\input{chapters/chapter5_implementation}

\input{chapters/chapter6_security}

\input{chapters/chapter7_discussion}

\input{chapters/chapter8_conclusions}

\bibliography{references}

\end{document}
